Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2021-07-27 edubots_surveys.Rnw
Changes during conference

The {\bf initial survey} was done in two separate Google forms, one for university and other for tertiary non-university teachers. Questions and responses were the same, except for the type of tertiary degrees that were taught. In both cases, we used mainly mailing lists, as well as Telegram groups, to spread the link to the form. The university form was spread in Spanish (mainly Andaluc\'{\i}a and Galicia) universities, and also Costa Rica and Mexico. The tertiary non-university teachers are based mainly in Andaluc\'{\i}a. Dissemination and answers took place in the first quarter of 2021, post-pandemic and while, at least in Spain, many universities had mandatory virtual teaching.

@@

% Describe the survey and population
We collected responses from 282 teachers from Spain and Spanish Speaking countries for the first survey (\textit{survey1} and 205 for the second one (\textit{survey2}). The two forms were open for approximately the same amount of time, around two months. The sample includes mostly teachers from university, although about 32\% and 8\% are from vocational education in the first and second survey respectively. Regarding gender, 61\% of the teachers were male and 32\% female, while approximately 5\% chose not to disclose it. Finally, the responses are more or less equally distributed according to the teaching experience, showing 24\% of responses from teachers with 5 or less years of experience, 26\% for 6-15 years of experience, 30\% for 16-25, and 20\% for teachers with more than 25 years of experience.

@@

Well try to find the answers to the four research questions next, by analyzing the response to the different survey questions.

@@

Overall, the majority of the teachers responded that they do use messaging apps in their classrooms, from which apps provided by the academic institution (n=159, 56\%) and WhatsApp (n=124, 44.0\%) were the most common responses. Only 19 teachers (6.7\%) replied that they do not use any messaging app in their class. %Remember to add something to the discussion below - JJ
%
\begin{table}[h!tbp]
\begin{tabular}{lllll}
\hline
\multicolumn{1}{c}{Messaging App} & \multicolumn{2}{c}{Yes} & \multicolumn{2}{l}{No} \\ \cline{2-5} 
\multicolumn{1}{c}{} & \multicolumn{1}{c}{Frequency} & \multicolumn{1}{c}{\%} & Frequency & \multicolumn{1}{c}{\%} \\ \hline
Telegram & 62 & 22.0 & 220 & 78.0 \\
WhatsApp & 124 & 44.0 & 158 & 56.0 \\
Slack & 15 & 5.3 & 267 & 94.7 \\
Other & 60 & 21.3 & 222 & 78.7 \\
\begin{tabular}[c]{@{}l@{}}Provided by the Academic\\ Institution\end{tabular} & 159 & 56.4 & 123 & 43.6 \\
None & 19 & 6.7 & 263 & 93.3 \\ \hline
\end{tabular}
\vspace{0.8em}
\caption{Use of messaging apps in class}
\label{tab:use_messapps}
\end{table}

% Fran: I think figures with data about the use of messaging apps should come here. Then, we say we also asked about changes after covid and include data about that
Results in Figs. \ref{figure:provided*chatbots_use}, \ref{figure:whatsapp*chatbots_use}, \ref{figure:telegram*chatbots_use} do not show significant differences in the use of instant messaging apps between teachers from University and vocational education. In general, most teachers prefer messaging apps provided by their own institutions (we also show the results for the two most popular messaging app platforms: Whatsapp and Telegram). With respect to specific disciplines, Engineering and Technology teachers are more active in their use, although it is also relevant the amount of teachers from Humanities that answered they used these apps in their classes (around 60\% use the apps provided by their institutions).

@@

One interesting result is that, contrary to the idea of technology adaptation issues when considering older teachers, about 65\% of teachers with more than 25 years of experience use the platforms provided by their institutions while the percentage goes down to less than 50\% for teachers with 6-15 years of experience.

@@

As it can be seen by the answer to the first and last question, in general teachers do not want to participate in a chat group with students; either they want to simply leave leave the students alone in their own chat group, or otherwise they prefer not to be part of that interaction. In general, that is going to be the case no matter what; it is well known that students organize their own chat groups with many (and not always conveyable) intentions, so teachers do not want to take any part in it. Overwhelmingly, they don't want to participate in any chat group with students, but even less so if it includes the rest of the faculty.

@@

% Emotinal, mood for future work... Chatbots perform sentiment analysis
Finally, the full extent of chatbot technology is not really being examined in these surveys. They can be connected to natural language processing engines with sentiment analysis as well as other analytics. Examining the mood of the class, as response to new material, assignment, exams or external events might help will help any student-centered teaching strategy, which will help students (and teachers) reach their learning objectives. This line of work, however, will be based on the introduction of some pilot study, combined with initial opinion assesment on the students.

>>>>

% Move a part of this to an appendix: Noorhan
The {\bf initial survey} was done in two separate Google forms, one for university and other for tertiary non-university teachers. Questions and responses were the same, except for the type of tertiary degrees that were taught. In both cases, we used mainly mailing lists, as well as Telegram groups, to spread the link to the form. The university form was spread in Spanish (mainly Andaluc\'{\i}a and Galicia) universities, and also Costa Rica and Mexico. The tertiary non-university teachers are based mainly in Andaluc\'{\i}a. Dissemination and answers took place in the first quarter of 2021, post-pandemic and while, at least in Spain, many universities had mandatory virtual teaching.

@@

% Describe the survey and population
We collected responses from 282 teachers from Spain and Spanish Speaking countries for \textit{survey1} and 205 for \textit{survey2}. The two forms were open for approximately the same amount of time, around two months. The sample includes mostly teachers from university, although about 32\% and 8\% are from vocational education in the first and second survey respectively. Regarding gender, 61\% of the teachers were male and 32\% female, while approximately 5\% chose not to disclose it. Finally, the responses are more or less equally distributed according to the teaching experience, showing 24\% of responses from teachers with 5 or less years of experience, 26\% for 6-15 years of experience, 30\% for 16-25, and 20\% for teachers with more than 25 years of experience.

@@

We will try to find the answers to the four research questions next, by analyzing the response to the different survey questions.

@@

Overall, the majority of the teachers responded that they do use messaging apps in their classrooms, from which apps provided by the academic institution (n=159, 56\%) and WhatsApp (n=124, 44.0\%) were the most common responses. Only 19 teachers (6.7\%) replied that they do not use any messaging app in their class. %Remember to add something to the discussion below - JJ
%
\begin{table}[h!tbp]
\begin{tabular}{lllll}
\hline
\multicolumn{1}{c}{Messaging App} & \multicolumn{2}{c}{Yes} & \multicolumn{2}{l}{No} \\ \cline{2-5} 
\multicolumn{1}{c}{} & \multicolumn{1}{c}{Frequency} & \multicolumn{1}{c}{\%} & Frequency & \multicolumn{1}{c}{\%} \\ \hline
Telegram & 62 & 22.0 & 220 & 78.0 \\
WhatsApp & 124 & 44.0 & 158 & 56.0 \\
Slack & 15 & 5.3 & 267 & 94.7 \\
Other & 60 & 21.3 & 222 & 78.7 \\
\begin{tabular}[c]{@{}l@{}}Provided by the Academic\\ Institution\end{tabular} & 159 & 56.4 & 123 & 43.6 \\
None & 19 & 6.7 & 263 & 93.3 \\ \hline
\end{tabular}
\vspace{0.8em}
\caption{Use of messaging apps to assist the learning process.}
\label{tab:use_messapps}
\end{table}

% Fran: I think figures with data about the use of messaging apps should come here. Then, we say we also asked about changes after covid and include data about that
Results in Figs. \ref{figure:provided*chatbots_use}, \ref{figure:whatsapp*chatbots_use}, \ref{figure:telegram*chatbots_use} do not show significant differences in the use of instant messaging apps between teachers from University and vocational education. In general, most teachers prefer messaging apps provided by their own institutions (we also show the results for the two most popular messaging app platforms: WhatsApp and Telegram). With respect to specific disciplines, Engineering and Technology teachers are more active in their use, although it is also relevant the amount of teachers from Humanities that answered they used these apps in their classes (around 60\% use the apps provided by their institutions).

@@

One interesting result is that about 65\% of teachers with more than 25 years of experience use the platforms provided by their institutions while the percentage goes down to less than 50\% for teachers with 6-15 years of experience.

@@

As it can be seen by the answer to the first and last question, in general teachers do not want to participate in a chat group with students; either they want to simply leave the students alone in their own chat group, or otherwise they prefer not to be part of that interaction. In general, that is going to be the case no matter what; it is well known that students organize their own chat groups with many (and not always conveyable) intentions, so teachers do not want to take any part in it. Overwhelmingly, they don't want to participate in any chat group with students, but even less so if it includes the rest of the faculty.

@@

% Emotinal, mood for future work... Chatbots perform sentiment analysis
Finally, the full extent of chatbot technology is not really being examined in these surveys. They can be connected to natural language processing engines with sentiment analysis as well as other analytics. Examining the mood of the class, as response to new material, assignment, exams or external events might help will help any student-centered teaching strategy, which will help students (and teachers) reach their learning objectives. This line of work, however, will be based on the introduction of some pilot study, combined with initial opinion assessment on the students.
